---
title: "Project MEDM"
output: html_document
        #pdf_document
---

<!-- Press Knit, to make a html-file of the code below. This we can use while working with the report to see how it is. Because the dataset flights in pretty large it takes some time to knit the file.. 
We don´t have to compilate the code to a pdf before we are delivering it, but if you want to you might have to download a program, maybe Latex or something, I am not quite sure. I already have it so I can compilate it to a pdf when we are done with it. 
But remember if you compilate it to a pdf-file some lines can be too long and get cutted. Then you have to go in the code and adjust it. If you need a linespace you need to have 3 space in the line above. -->

<!-- In gitkranken: 
When you have edited the code, save it usual. Then open gitkranken, and you will see a blue line with 
" x file changes in working directory" and a small box "View changes" in the above right corner. Click this one. First you have to commit what you have done, i.e. click on the Project_report.html-file on the right side and press Stage File. (And don´t stage all the other files) Then make a comment on what you have changed in the file, in the bottom right corner, and press the green box "Stage files/changes to commit". 
After commiting you first have to pull what the other have done, i.e. press "Pull". If you get "pulled successfully" down in the left corner everything is okay and you can press "Push". 
If not, then we might have some merge conflict we need to fix. 
Always remember to do all these steps
1. Commit
2. Pull
3. Push

If you have edited the file, but don´t want to save it, only download what the others have done, you can press the "Stash" instead of pull and push. 
-->

```{r setup, include=FALSE}
#Have to include the files in the same folder the Rmarkdown is in.
airlines = read.csv("airlines.csv")
airports = read.csv("airports.csv")
flights = read.csv("flights.csv")

```

   
```{r echo = FALSE}  
 #to prevent the code chunk from printing of the R code

```

<!-- This is a comment. -->


Here we can write Latex code, with formulas inline $E=mc^2$, or in a new line,
$$
    A = \pi \cdot r^2.
$$
This is standard Latex-code, so we can write all mathematical signs by using $\mathbf{A}x = b$. 
$\hat{Y}$, $Y_i$, $\theta$, $\rho(X,C)$, $x_1^{(1)}$. 
\textbf{bold text}, \textit{italicized}
This link includes a list of mathematical symbols written in latex. https://oeis.org/wiki/List_of_LaTeX_mathematical_symbols


   <!-- 1 # means Section  -->
# PART 1: EXPLORATORTY DATA ANALYSIS 

 <!-- Perform a exploratory data analysis and discuss what you have learned from this analysis.
 
 (Just testing the app)
 
 Aim: Which airline should we fly on? 
 
The U.S. Department of Transportation's (DOT) Bureau of Transportation Statistics tracks the on-time performance of domestic flights operated by large air carriers. 
Summary information on the number of on-time, delayed, canceled, and diverted flights is published in DOT's monthly Air Travel Consumer Report and in this dataset of 2015 flight delays and cancellations.

Initial steps: 
1-	To solve the classification problem the group must create a new variable through the analysis of the variables associated to the delay that can be used as response variable for the problem aim. 
 
 -->


```{r , echo=TRUE}
library(ggplot2)  #for ggplot
library(class)    #for confusion matrices
library(caret)    #for knn function

```



# PART 2: SOLVE CLASSIFICATION PROBLEM

 <!-- Solve your classification problem.  Consider several classification methods and discuss  how  can  they  contribute  for  the  solution  of  your  problem.   Include in your discussion topics such as the options that you have made in building each  classifier;  interpretation  of  results;  validation  of  the  methods  used  and possible assumptions;  advantages and disadvantages of each alternative;  etc. Have  in  mind  that  some  of  the  explanatory  variables  may  be  irrelevant  to the  classification problem and that  you may need to  do  some  preprocessing methodologies of your data set e.g.  dimensionality reduction techniques.-->

```{r , echo=TRUE}


```
 <!-- 2 ## means subsection, 3 means subsubsection etc.-->   
   
## Resubstitution

```{r , echo=TRUE}


```


## Hold Out method

```{r , echo=TRUE}


```


## LOOCV

```{r , echo=TRUE}


```


## K-fold CV

```{r , echo=TRUE}


```


## Repeated K-fold CV

```{r , echo=TRUE}


```



## Bootstrap

```{r , echo=TRUE}


```



# PART 3: LEARNING/LIMITATIONS/FUTURE WORK

 <!-- Imagine you are going to meet the researcher who contacted you.  Report to him/her what you have learned about the problem.  Discuss limitations of the analysis you have done and provide suggestions for future work.-->


```{r , echo=TRUE}


```

