R_project - MEDM

---
title: "Project MEDM"
output: html_document
        #pdf_document
        
---

#Load data set
```{r setup, include=FALSE}
#Have to include the files in the same folder the Rmarkdown is in.
airlines = read.csv("airlines.csv")
airports = read.csv("airports.csv")
flights = read.csv("flights.csv")
```

# Run Libraries
```{r , echo=FALSE}
#Libraries
library(ggplot2)       #for ggplot
library(class)         #for confusion matrices
library(caret)         #for knn function
library(kknn)
library(e1071)         #for naive bayes function
library(GGally)
library(corrplot)
library(DataExplorer)
library(dplyr)
# library(MASS)          #for LDA
library(nnet)         
# library(leaps)         #for Forward selection
library(colorspace)    #for plot colors
library(dplyr)         #for barplots
library(tidyverse)     #for barplots
# library(StatMatch)     #for knngow function
# library(dprep)        #for knngow function
library(broom)
library(rpart)        #for fitting tree
library(rpart.plot)   #for plotting tree
library(leaflet)      #for map plots
library(randomForest) # for randomForest
library(plotrix)
```


# Introduction to data set
```{r , echo=TRUE}
handled_flights <- flights[,] 

##Pie chart for diverted/cancelled/none
count_diverted <- sum(handled_flights$DIVERTED==1)
count_cancelled <- sum(handled_flights$CANCELLED==1)
count_scheduled <- sum(handled_flights$DIVERTED==0 & handled_flights$CANCELLED==0)

slices <- c(count_diverted, count_cancelled, count_scheduled) 
lbls <- c("Diverted", "Cancelled", "Scheduled")
pct <- round(slices/sum(slices)*100, 2)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # add % to labels 
pie(slices,labels = lbls, col=c("firebrick","brown4","chartreuse3"),   #Not very good - use Carolina's chart
   main="Percentage of flight types") 
```

```{r , echo=TRUE}
#Erase observations correspondent to cancelled and diverted flights
handled_flights <- handled_flights[handled_flights$DIVERTED==0,]
handled_flights <- handled_flights[handled_flights$CANCELLED==0,]

#Ensure that we are only dealing with observations for which Arrival delay is defined
handled_flights<- handled_flights[is.na(handled_flights$ARRIVAL_DELAY)==FALSE,]

# Ensure that "" factor for FLIGHT_NUMBER is inexistent
handled_flights<- droplevels(handled_flights,exclude = handled_flights$TAIL_NUMBER=="")

#Erase columns which are not relevant to our problem
handled_flights$DIVERTED <- NULL
handled_flights$CANCELLED <- NULL
handled_flights$CANCELLATION_REASON <- NULL
```

```{r , echo=TRUE}
#Change handled_flights dataframe data types to the correct ones
#Discrete variables must be considered as FACTORs, continuous variables should be considered as NUMERICs
handled_flights$YEAR <- as.factor(handled_flights$YEAR)
counts_year <- as.matrix(table(handled_flights$YEAR)) 
counts_year_mean <- colSums(counts_year)/dim(counts_year)[1]
max(counts_year)
min(counts_year)
median(counts_year)
sd(counts_year)

handled_flights$MONTH <- as.factor(handled_flights$MONTH)
counts_month <- as.matrix(table(handled_flights$MONTH)) 
counts_month_mean <- colSums(counts_month)/dim(counts_month)[1]
max(counts_month)
min(counts_month)
median(counts_month)
sd(counts_month)


handled_flights$DAY <- as.factor(handled_flights$DAY)
counts_day <- as.matrix(table(handled_flights$DAY)) 
counts_day_mean <- colSums(counts_day)/dim(counts_day)[1]
max(counts_day)
min(counts_day)
median(counts_day)
sd(counts_day)

handled_flights$DAY_OF_WEEK <- as.factor(handled_flights$DAY_OF_WEEK)
counts_dayofweek <- as.matrix(table(handled_flights$DAY_OF_WEEK)) 
counts_dayofweek_mean <- colSums(counts_dayofweek)/dim(counts_dayofweek)[1]
max(counts_dayofweek)
min(counts_dayofweek)
median(counts_dayofweek)
sd(counts_dayofweek)

counts_airline <- as.matrix(table(handled_flights$AIRLINE))
counts_airline_mean <- colSums(counts_airline)/dim(counts_airline)[1]
max(counts_airline)
min(counts_airline)
median(counts_airline)
sd(counts_airline)


handled_flights$FLIGHT_NUMBER <- as.factor(handled_flights$FLIGHT_NUMBER)
counts_flightnumber <- as.matrix(table(handled_flights$FLIGHT_NUMBER))
counts_flightnumber_mean <- colSums(counts_flightnumber)/dim(counts_flightnumber)[1]
max(counts_flightnumber)
min(counts_flightnumber)
median(counts_flightnumber)
sd(counts_flightnumber)

counts_tailnumber <- as.matrix(table(handled_flights$TAIL_NUMBER))
counts_tailnumber_mean <- colSums(counts_tailnumber)/dim(counts_tailnumber)[1]
max(counts_tailnumber)
min(counts_tailnumber)
median(counts_tailnumber)
sd(counts_tailnumber)

counts_origin_airport <- as.matrix(table(handled_flights$ORIGIN_AIRPORT))
counts_origin_airport_mean <- colSums(counts_origin_airport)/dim(counts_origin_airport)[1]
max(counts_origin_airport)
min(counts_origin_airport)
median(counts_origin_airport)
sd(counts_origin_airport)

counts_destination_airport <- as.matrix(table(handled_flights$DESTINATION_AIRPORT))
counts_destination_airport_mean <- colSums(counts_destination_airport)/dim(counts_destination_airport)[1]
max(counts_destination_airport)
min(counts_destination_airport)
median(counts_destination_airport)
sd(counts_destination_airport)




handled_flights$SCHEDULED_DEPARTURE <- as.numeric(handled_flights$SCHEDULED_DEPARTURE)
min(handled_flights$SCHEDULED_DEPARTURE)
max(handled_flights$SCHEDULED_DEPARTURE)
mean(handled_flights$SCHEDULED_DEPARTURE)
median(handled_flights$SCHEDULED_DEPARTURE)
sd(handled_flights$SCHEDULED_DEPARTURE)

handled_flights$DEPARTURE_TIME <- as.numeric(handled_flights$DEPARTURE_TIME)
min(handled_flights$DEPARTURE_TIME)
max(handled_flights$DEPARTURE_TIME)
mean(handled_flights$DEPARTURE_TIME)
median(handled_flights$DEPARTURE_TIME)
sd(handled_flights$DEPARTURE_TIME)

handled_flights$DEPARTURE_DELAY <- as.numeric(handled_flights$DEPARTURE_DELAY)
min(handled_flights$DEPARTURE_DELAY)
max(handled_flights$DEPARTURE_DELAY)
mean(handled_flights$DEPARTURE_DELAY)
median(handled_flights$DEPARTURE_DELAY)
sd(handled_flights$DEPARTURE_DELAY)

handled_flights$TAXI_OUT <- as.numeric(handled_flights$TAXI_OUT)
min(handled_flights$TAXI_OUT)
max(handled_flights$TAXI_OUT)
mean(handled_flights$TAXI_OUT)
median(handled_flights$TAXI_OUT)
sd(handled_flights$TAXI_OUT)

handled_flights$WHEELS_OFF <- as.numeric(handled_flights$WHEELS_OFF)
min(handled_flights$WHEELS_OFF)
max(handled_flights$WHEELS_OFF)
mean(handled_flights$WHEELS_OFF)
median(handled_flights$WHEELS_OFF)
sd(handled_flights$WHEELS_OFF)

handled_flights$SCHEDULED_TIME <- as.numeric(handled_flights$SCHEDULED_TIME)
min(handled_flights$SCHEDULED_TIME)
max(handled_flights$SCHEDULED_TIME)
mean(handled_flights$SCHEDULED_TIME)
median(handled_flights$SCHEDULED_TIME)
sd(handled_flights$SCHEDULED_TIME)

handled_flights$ELAPSED_TIME <- as.numeric(handled_flights$ELAPSED_TIME)
min(handled_flights$ELAPSED_TIME)
max(handled_flights$ELAPSED_TIME)
mean(handled_flights$ELAPSED_TIME)
median(handled_flights$ELAPSED_TIME)
sd(handled_flights$ELAPSED_TIME)

handled_flights$AIR_TIME <- as.numeric(handled_flights$AIR_TIME)
min(handled_flights$AIR_TIME)
max(handled_flights$AIR_TIME)
mean(handled_flights$AIR_TIME)
median(handled_flights$AIR_TIME)
sd(handled_flights$AIR_TIME)

handled_flights$DISTANCE <- as.numeric(handled_flights$DISTANCE)
min(handled_flights$DISTANCE)
max(handled_flights$DISTANCE)
mean(handled_flights$DISTANCE)
median(handled_flights$DISTANCE)
sd(handled_flights$DISTANCE)

handled_flights$WHEELS_ON <- as.numeric(handled_flights$WHEELS_ON)
min(handled_flights$WHEELS_ON)
max(handled_flights$WHEELS_ON)
mean(handled_flights$WHEELS_ON)
median(handled_flights$WHEELS_ON)
sd(handled_flights$WHEELS_ON)

handled_flights$TAXI_IN <- as.numeric(handled_flights$TAXI_IN)
min(handled_flights$TAXI_IN)
max(handled_flights$TAXI_IN)
mean(handled_flights$TAXI_IN)
median(handled_flights$TAXI_IN)
sd(handled_flights$TAXI_IN)

handled_flights$SCHEDULED_ARRIVAL <- as.numeric(handled_flights$SCHEDULED_ARRIVAL)
min(handled_flights$SCHEDULED_ARRIVAL)
max(handled_flights$SCHEDULED_ARRIVAL)
mean(handled_flights$SCHEDULED_ARRIVAL)
median(handled_flights$SCHEDULED_ARRIVAL)
sd(handled_flights$SCHEDULED_ARRIVAL)

handled_flights$ARRIVAL_TIME <- as.numeric(handled_flights$ARRIVAL_TIME)
min(handled_flights$ARRIVAL_TIME)
max(handled_flights$ARRIVAL_TIME)
mean(handled_flights$ARRIVAL_TIME)
median(handled_flights$ARRIVAL_TIME)
sd(handled_flights$ARRIVAL_TIME)

handled_flights$ARRIVAL_DELAY <- as.numeric(handled_flights$ARRIVAL_DELAY)
min(handled_flights$ARRIVAL_DELAY)
max(handled_flights$ARRIVAL_DELAY)
mean(handled_flights$ARRIVAL_DELAY)
median(handled_flights$ARRIVAL_DELAY)
sd(handled_flights$ARRIVAL_DELAY)
```

# Response variable

```{r , echo=TRUE}
#Define 2 categories for the response variable (None, Delayed) 
handled_flights$DELAY <- as.factor( ifelse(handled_flights$ARRIVAL_DELAY < 1,handled_flights$DELAY <- "None", handled_flights$DELAY <- "Delayed"))

#Check response variable categories distribution in handled_flights
count_delays <- as.matrix(table(handled_flights$DELAY))
count_delayed <-count_delays[1]
count_nodelayed <-count_delays[2]


slices <- c(count_delayed, count_nodelayed) 
lbls <- c("Delayed", "None")
pct <- round(slices/sum(slices)*100, 2)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # add % to labels 
pie(slices,labels = lbls, col=c("firebrick2", "green3"),   
   main="Distribution of response variable classes") 

#CONSIDER 4 CATEGORIES FOR RESPONSE ( We tried... :( )
    #To define the degree of delay, we consider the median and mean value, for example, considering only the         values   of ARRIVAL_DELAY > 0, which represent delays
    #median(handled_flights$ARRIVAL_DELAY[handled_flights$ARRIVAL_DELAY>0]) #mean = 33.11 median = 15  
    #mean(handled_flights$ARRIVAL_DELAY[handled_flights$ARRIVAL_DELAY>0])

    # handled_flights$DELAY <- as.factor( ifelse(handled_flights$ARRIVAL_DELAY < 1,handled_flights$DELAY <-           "None", ifelse(handled_flights$ARRIVAL_DELAY <16,handled_flights$DELAY <- "Low",                                 ifelse(handled_flights$ARRIVAL_DELAY <34,handled_flights$DELAY <- "Medium", handled_flights$DELAY <-             "High"))) )
    # 
  
    #Separate data into the 4 categories and report results
    #nodelay <- flights_to_predict[flights_to_predict$DELAY=="None",]
    #lowdelay <- flights_to_predict[flights_to_predict$DELAY=='Low',]
    #mediumdelay <- flights_to_predict[flights_to_predict$DELAY=='Medium',]
    #highdelay <- flights_to_predict[flights_to_predict$DELAY=='High',]
    
    # create_report(nodelay, output_file = 'nodelay.html')
    # create_report(lowdelay, output_file = 'lowdelay.html')
    # create_report(mediumdelay, output_file = 'mediumdelay.html')
    # create_report(highdelay, output_file = 'highdelay.html')


#Since our objective is to predict the delay of a flight, we must give to the classifier only the variables      corresponding to subjects known a priori

#Maintain a priori columns and remove a posteriori columns
handled_flights$DEPARTURE_TIME <- NULL
handled_flights$DEPARTURE_DELAY <- NULL
handled_flights$TAXI_OUT <- NULL                             
handled_flights$WHEELS_OFF <- NULL
handled_flights$ELAPSED_TIME <- NULL
handled_flights$AIR_TIME <- NULL
handled_flights$DISTANCE <- NULL    
handled_flights$WHEELS_ON <- NULL
handled_flights$TAXI_IN <- NULL
handled_flights$ARRIVAL_TIME <- NULL
handled_flights$ARRIVAL_DELAY <- NULL
handled_flights$AIR_SYSTEM_DELAY <- NULL
handled_flights$SECURITY_DELAY <- NULL
handled_flights$AIRLINE_DELAY <- NULL
handled_flights$LATE_AIRCRAFT_DELAY <- NULL
handled_flights$WEATHER_DELAY <- NULL
```

# Data exploration
```{r , echo=TRUE}
#Remove columns YEAR (2015 for all observations)
handled_flights$YEAR <- NULL

#Create list of existent airports
airports_list <- airports$IATA_CODE

#Create list of existent airlines
airlines_list <- airlines$IATA_CODE

#Show airports with invalid IATA codes
ghost_airports <- handled_flights[!(handled_flights$ORIGIN_AIRPORT %in% airports_list),]
ghost_airports<- droplevels(ghost_airports,exclude = ghost_airports$ORIGIN_AIRPORT %in% airports_list)

ghost_airports <- ghost_airports[!(ghost_airports$DESTINATION_AIRPORT %in% airports_list),]
ghost_airports<- droplevels(ghost_airports,exclude = ghost_airports$DESTINATION_AIRPORT %in% airports_list)

table(ghost_airports$MONTH) #all the flights are from october

barplot(table(ghost_airports$ORIGIN_AIRPORT), main="Frequencies of ghost flights by Origin Airport",
xlab="Origin Airport", xaxt = "n", col=c("gray0"), beside=TRUE)

barplot(table(ghost_airports$DESTINATION_AIRPORT), main="Frequencies of ghost flights by Destination Airport",
xlab="Destination Airport", xaxt = "n", col=c("gray0"), beside=TRUE)




#Erase observations for origin_airports, destination_airports which are not contained in the data frame airports and airlines which are not contained in the data frame airlines
handled_flights<- handled_flights[handled_flights$ORIGIN_AIRPORT %in% airports_list,]
handled_flights<- droplevels(handled_flights,exclude = !(handled_flights$ORIGIN_AIRPORT %in% airports_list))

handled_flights<- handled_flights[handled_flights$DESTINATION_AIRPORT %in% airports_list,]
handled_flights<- droplevels(handled_flights,exclude = !(handled_flights$DESTINATION_AIRPORT %in% airports_list))

handled_flights<- handled_flights[handled_flights$AIRLINE %in% airlines_list,]
handled_flights<- droplevels(handled_flights,exclude = !(handled_flights$AIRLINE %in% airlines_list))

barplot(table(handled_flights$ORIGIN_AIRPORT), main="Frequencies of valid flights by Origin Airport",
xlab="Origin Airport", xaxt = "n", col=c("gray0"), beside=TRUE)

barplot(table(handled_flights$DESTINATION_AIRPORT), main="Frequencies of valid flights by Destination Airport",
xlab="Destination Airport", xaxt = "n", col=c("gray0"), beside=TRUE)



#create_report(handled_flights, output_file = 'report_corr.html')

#Verify that flight number and tail number have no influence

#Tables with only non delayed/delayed flights
non_delayed <- handled_flights[handled_flights$DELAY=='None',]
delayed <-  handled_flights[handled_flights$DELAY=='Delayed',]

barplot(table(non_delayed$FLIGHT_NUMBER), main="Frequencies of non delayed flights by Flight Number",
  xlab="Flight Number", xaxt = "n",col=c("chartreuse3"), beside=TRUE)

barplot(table(delayed$FLIGHT_NUMBER), main="Frequencies of delayed flights by Flight Number",
  xlab="Flight Number", xaxt = "n",col=c("firebrick"), beside=TRUE)

barplot(table(non_delayed$TAIL_NUMBER), main="Frequencies of non delayed flights by Tail Number",
  xlab="Tail Number", xaxt = "n",col=c("chartreuse3"), beside=TRUE)

barplot(table(delayed$TAIL_NUMBER), main="Frequencies of delayed flights by Tail Number",
  xlab="Tail Number", xaxt = "n", col=c("firebrick"), beside=TRUE)
```

```{r , echo=TRUE}
#Erase columns with too many categories
handled_flights$FLIGHT_NUMBER <- NULL
handled_flights$TAIL_NUMBER <- NULL
```

```{r , echo=TRUE}
#Barplot of the delays per Month
#Total
barplot(table(handled_flights$MONTH), main="Frequencies of flights by Month",
  xlab="Month", col=c("darkturquoise"), args.legend = list(x="bottomleft"),
  legend = rownames(handled_flights$MONTH), beside=TRUE)

barplot(table(handled_flights$DELAY, handled_flights$MONTH), main="Distribution of response variable classes by Month",
  xlab="Month", col=c("firebrick", "chartreuse3"), args.legend = list(x="bottomleft"),
  legend = rownames(table(handled_flights$DELAY, handled_flights$MONTH)), beside=TRUE)

#Barplot of the delays per Day
#Total
barplot(table(handled_flights$DAY), main="Frequencies of flights by Day",
  xlab="Day", col=c("darkturquoise"), args.legend = list(x="bottomleft"),
  legend = rownames(handled_flights$DAY), beside=TRUE)

barplot(table(handled_flights$DELAY, handled_flights$DAY), main="Distribution of response variable classes by Day",
  xlab="Day", col=c("firebrick", "chartreuse3"), args.legend = list(x="bottomleft"),
  legend = rownames(table(handled_flights$DELAY, handled_flights$DAY)), beside=TRUE)

#Barplot of the delays per Day Of Week
#Total
barplot(table(handled_flights$DAY_OF_WEEK), main="Frequencies of flights by Day Of Week",
  xlab="Day Of Week", col=c("darkturquoise"), args.legend = list(x="bottomleft"),
  legend = rownames(handled_flights$DAY_OF_WEEK), beside=TRUE)

barplot(table(handled_flights$DELAY, handled_flights$DAY_OF_WEEK), main="Distribution of response variable classes by Day Of Week",
  xlab="Day Of Week", col=c("firebrick", "chartreuse3"), args.legend = list(x="bottomleft"),
  legend = rownames(table(handled_flights$DELAY, handled_flights$DAY_OF_WEEK)), beside=TRUE)

#Barplot of the delays per Airline 
#Total
barplot(table(handled_flights$AIRLINE), main="Frequencies of flights by Airline",
  xlab="Airline", col=c("darkturquoise"), args.legend = list(x="bottomleft"),
  legend = rownames(handled_flights$AIRLINE), beside=TRUE)

barplot(table(handled_flights$DELAY, handled_flights$AIRLINE), main="Distribution of response variable classes by Airline",
  xlab="Airlines", col=c("firebrick", "chartreuse3"), args.legend = list(x="topleft"),
  legend = rownames(table(handled_flights$DELAY, handled_flights$AIRLINE)), beside=TRUE)


#Barplot of the delays per Scheduled Departure
#Total
# set up cut-off values 
breaks <- c(0:24)*100
# specify interval/bin labels
# tags <- c("[0-1)","[1-2)", "[2-3)", "[3-4)", "[4-5)", "[5-6)","[6-7)", "[7-8)","[8-9)", "[9-10)", "[10-11)",
#           "[11-12)","[12-13)", "[13-14)", "[14-15)", "[15-16)", "[16-17)","[17-18)", "[18-19)","[19-20)", "[20-21)", "[21-22)", "[22-23)","[23-24)")
tags <- c("0","1", "2",  "3", "4", "5",  "6",  "7",  "8",  "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", "21", "22", "23")
#bucketing values into bins
group_tags_departure <- cut(handled_flights$SCHEDULED_DEPARTURE, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags) 

barplot(table(group_tags_departure), main="Distribution of flights by Scheduled Departure Hour",
  xlab="Scheduled Departure Hour", col=c("darkturquoise"), cex.names=0.7, beside=TRUE)


#By Category ------------------------------------------------------------------------------
#Total
# set up cut-off values 
breaks <- c(0:24)*100
# specify interval/bin labels
tags <- c("0","1", "2",  "3", "4", "5",  "6",  "7",  "8",  "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", "21", "22", "23")
group_tags_departure_nondelayed <- cut(non_delayed$SCHEDULED_DEPARTURE,  
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)

group_tags_departure_delayed <- cut(delayed$SCHEDULED_DEPARTURE,  
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)

table_departure <- rbind(table(group_tags_departure_delayed), table(group_tags_departure_nondelayed))
row.names(table_departure) <- c("Delayed","None")


barplot(table_departure, main="Distribution of response variable classes by Scheduled Departure Hour",
  xlab="Scheduled Departure Hour", col=c("firebrick", "chartreuse3"), args.legend = list(x="topleft"),
  legend = c("Delayed","None"), cex.names=0.7, beside=TRUE)



#Barplot of the delays per Scheduled Arrival
# bucketing values into bins
group_tags_arrival <- cut(handled_flights$SCHEDULED_ARRIVAL, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)

barplot(table(group_tags_arrival), main="Distribution of flights by Scheduled Arrival Hour",
  xlab="Scheduled Arrival Hour", col=c("darkturquoise"), cex.names=0.7, beside=TRUE)


#By Category  ------------------------------------------------------------------------------
group_tags_arrival_nondelayed <- cut(non_delayed$SCHEDULED_ARRIVAL,  
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)

group_tags_arrival_delayed <- cut(delayed$SCHEDULED_ARRIVAL,  
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)

table_arrival <- rbind(table(group_tags_arrival_delayed), table(group_tags_arrival_nondelayed))
row.names(table_arrival) <- c("Delayed","None")


barplot(table_arrival, main="Distribution of response variable classes by Scheduled Arrival Hour",
  xlab="Scheduled Arrival Hour", col=c("firebrick", "chartreuse3"), args.legend = list(x="topleft"),
  legend = c("Delayed","None"), cex.names=0.7, beside=TRUE)





#Barplot of the delays per Scheduled Time
breaks <- c(0:15)*50
# tags <- c("[0-50)","[50-100)", "[100-150)", "[150-200)", "[200-250)", "[250-300)","[300-350)", "[350-400)","[400-450)", "[450-500)", "[500-550)", "[550-600)","[600-650)", "[650-700)", "[700-750)")
tags <- c("0","50", "100", "150", "200", "250","300", "350","400", "450", "500", "550","600", "650", "700")

group_tags_time <- cut(handled_flights$SCHEDULED_TIME, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)

barplot(table(group_tags_time), main="Distribution of flights by Scheduled Time",
  xlab="Scheduled Time (minutes)", col=c("darkturquoise"), cex.names=0.7, beside=TRUE)


#By Category  ------------------------------------------------------------------------
group_tags_time_nondelayed <- cut(non_delayed$SCHEDULED_TIME,  
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)

group_tags_time_delayed <- cut(delayed$SCHEDULED_TIME,  
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)

table_time <- rbind(table(group_tags_time_delayed), table(group_tags_time_nondelayed))
row.names(table_time) <- c("Delayed","None")


barplot(table_time, main="Distribution of response variable classes by Scheduled Time",
  xlab="Scheduled Time (minutes)", col=c("firebrick", "chartreuse3"), args.legend = list(x="topright"),
  legend = c("Delayed","None"), cex.names=0.7, beside=TRUE)



#Schedule Departure Metrics
mean(non_delayed$SCHEDULED_DEPARTURE)
min(non_delayed$SCHEDULED_DEPARTURE)
max(non_delayed$SCHEDULED_DEPARTURE)
median(non_delayed$SCHEDULED_DEPARTURE)
sd(non_delayed$SCHEDULED_DEPARTURE)

mean(delayed$SCHEDULED_DEPARTURE)
min(delayed$SCHEDULED_DEPARTURE)
max(delayed$SCHEDULED_DEPARTURE)
median(delayed$SCHEDULED_DEPARTURE)
sd(delayed$SCHEDULED_DEPARTURE)

#Schedule Arrival Metrics
mean(non_delayed$SCHEDULED_ARRIVAL)
min(non_delayed$SCHEDULED_ARRIVAL)
max(non_delayed$SCHEDULED_ARRIVAL)
median(non_delayed$SCHEDULED_ARRIVAL)
sd(non_delayed$SCHEDULED_ARRIVAL)

mean(delayed$SCHEDULED_ARRIVAL)
min(delayed$SCHEDULED_ARRIVAL)
max(delayed$SCHEDULED_ARRIVAL)
median(delayed$SCHEDULED_ARRIVAL)
sd(delayed$SCHEDULED_ARRIVAL)

#Schedule Time Metrics
mean(non_delayed$SCHEDULED_TIME)
min(non_delayed$SCHEDULED_TIME)
max(non_delayed$SCHEDULED_TIME)
median(non_delayed$SCHEDULED_TIME)
sd(non_delayed$SCHEDULED_TIME)

mean(delayed$SCHEDULED_TIME)
min(delayed$SCHEDULED_TIME)
max(delayed$SCHEDULED_TIME)
median(delayed$SCHEDULED_TIME)
sd(delayed$SCHEDULED_TIME)

```

## Final data set

```{r , echo=TRUE}
set.seed(0)
flights_small <- sample_n(handled_flights, 30000)   

#Plot distribution of categories by Airline in flights_small
barplot(table(flights_small$DELAY, flights_small$AIRLINE), main="Distribution of response variable classes by Airline",
  xlab="Airlines", col=c("firebrick", "chartreuse3"), args.legend = list(x="topleft"),
  legend = rownames(table(flights_small$DELAY, flights_small$AIRLINE)), beside=TRUE)


#Guarantee sampling distributions are OK

#Find the percentages of each airline
n = dim(handled_flights)[1]
counts <- table(handled_flights$AIRLINE)
percentage <- counts/n
percentage

#Find the percentages of each airline in small dataset
n2 = dim(flights_small)[1]
counts2 <- table(flights_small$AIRLINE)
percentage2 <- counts2/n2
percentage2
#Comment that we see that these percentages is the same so this is okay to do

#Count by airline how many flights are delayed in handled_flights
handled_flightsdelayed<- handled_flights[handled_flights$ARRIVAL_DELAY>0,]
counts3 <- table(handled_flightsdelayed$AIRLINE)
percentdelayedbyairline <- counts3/counts

#Count by airline how many flights are delayed in flights_small
flights_smalldelayed<- flights_small[flights_small$ARRIVAL_DELAY>0,]
counts4 <- table(flights_smalldelayed$AIRLINE)
percentdelayedbyairline2 <- counts4/counts2

percentdelayedbyairline
percentdelayedbyairline2 

#Check if the variable distributions are maintained after sampling
# create_report(flights_small, output_file = 'report_flights_small.html')
# create_report(handled_flights, output_file = 'report_handled_flights.html')

```

# PART 2: SOLVE CLASSIFICATION PROBLEM

 <!-- Solve your classification problem.  Consider several classification methods and discuss  how  can  they  contribute  for  the  solution  of  your  problem.   Include in your discussion topics such as the options that you have made in building each  classifier;  interpretation  of  results;  validation  of  the  methods  used  and possible assumptions;  advantages and disadvantages of each alternative;  etc. Have  in  mind  that  some  of  the  explanatory  variables  may  be  irrelevant  to the  classification problem and that  you may need to  do  some  preprocessing methodologies of your data set e.g.  dimensionality reduction techniques.-->

## Hold Out Method
```{r , echo=TRUE}
#Split sample in train (70%), validation (15%) and test (15%)
#Train: create the model
#Validation: choose parameters for classifiers
#Test: evaluate the fit of the classifier and compare it with the others

set.seed(0)   # To fix the random process at splitting data
num_obs <- dim(flights_small)[1]
index_list <- c(1:num_obs)
train_index <- sample(index_list, num_obs*0.7)
test_valid_index <- setdiff(index_list, train_index)
test_index <- sample(test_valid_index, length(test_valid_index)*0.5)
valid_index <- setdiff(test_valid_index, test_index)

train_data <- flights_small[train_index,]
train_data_x <- flights_small[train_index,]
train_data_x$DELAY <- NULL 
train_data_y <- (flights_small[train_index,])$DELAY


valid_data <- flights_small[valid_index,]
valid_data_x <- flights_small[valid_index,]
valid_data_x$DELAY <- NULL 
valid_data_y <- (flights_small[valid_index,])$DELAY

#Check percentages of delay and none in valid set
percent <- table(valid_data$DELAY)/dim(valid_data)[1]


test_data <- flights_small[test_index,]
test_data_x <- flights_small[test_index,]
test_data_x$DELAY <- NULL 
test_data_y <- (flights_small[test_index,])$DELAY

#Checking distribution in original and partitioned data  
# prop.table(table(train_data_y)) * 100 # at training set 
# prop.table(table(test_data_y)) * 100  # at testing set
# prop.table(table(valid_data_y)) * 100  # at validation set
# prop.table(table(flights_small$DELAY)) * 100  # at original dataset
#THIS SEEMS FINE!
```

## Naive Bayes method

```{r , echo=TRUE}
set.seed(0)

#Use validation set to determine if is better to use Laplace or not
#No Laplace
modelNBNoLap <- naiveBayes(formula = formula(train_data_y~.), data = train_data, laplace = 0)
predictNBNoLap <- predict(modelNBNoLap, valid_data_x)

#Evaluation
confusionMatrix(data = predictNBNoLap, reference = valid_data_y, positive = "None")

#With Laplace
modelNBLap <- naiveBayes(formula = formula(train_data_y~.), data = train_data, laplace = 1)
predictNBLap <- predict(modelNBLap, valid_data_x)

#Evaluation
confusionMatrix(data = predictNBLap, reference = valid_data_y, positive = "None")


#SO, WE DO NOT USE LAPLACE

predictNB <- predict(modelNBNoLap, test_data_x)

#Evaluation
confusionMatrix(data = predictNB, reference = test_data_y, positive = "None")


#BACKWARD PROBLEM: FIND THE AIRLINE ------------------------------- 

#Find the best airline using the histogram
tableNB <- cbind(test_data_x, predictNB)


#Barplot of the delays per Airline  
barplot(table(tableNB$predictNB, tableNB$AIRLINE), main="Categories of delay by Airline",
  xlab="Airlines", col=c("firebrick", "chartreuse3"), args.legend = list(x="topleft"),
  legend = rownames(table(tableNB$predictNB, tableNB$AIRLINE)), beside=TRUE)

#Get the percentages of delayed and not delayed for Airline
countdelayNB <- as.matrix(table(tableNB$predictNB, tableNB$AIRLINE))
countdelayNB <- sweep(countdelayNB, 2, colSums(countdelayNB), FUN = '/')

colnames(countdelayNB) = c("AA","AS", "B6","DL","EV","F9","HA","MQ","NK","OO","UA","US","VX","WN") 
rownames(countdelayNB) = c("Delayed", "None")

#Get the most and least delayed Airline
mostdelayedNB <- colnames(countdelayNB)[which.is.max(countdelayNB[1,])] 
mostdelayedNB      #Airline which had the higher value of delay

leastdelayedNB <- colnames(countdelayNB)[which.is.max(countdelayNB[2,])] 
leastdelayedNB      #Airline which had the lower value of delay

#See sorted list of delayed and non delayed
mostdelayed_sortNB <- sort(countdelayNB[1,], decreasing = TRUE) #Higher percentage means high delay
leastdelayed_sortNB <- sort(countdelayNB[2,], decreasing = TRUE) #Higher percentage means low delay

mostdelayed_sortNB

#Barplot of the delays (percentge) per Airline  
barplot(countdelayNB, main="Categories of delay by Airline - Percentage",
  xlab="Airlines", col=c("firebrick", "chartreuse3"), args.legend = list(x="topright"),
  legend = rownames(table(tableNB$predictNB, tableNB$AIRLINE)), beside=TRUE)
```


## Decision Tree
```{r , echo=TRUE}
set.seed(10)
#Use validation set to choose the splitting criterion (information or gini)
# MODEL FITTING

modelTREEgini <- rpart(formula = formula(train_data_y~.), data = train_data_x, method = "class", parms = list(split ="gini")) 

modelTREEinf <- rpart(formula = formula(train_data_y~.), data = train_data_x, method = "class", parms = list(split ="information")) 


predictTREEgini <- predict(modelTREEgini, valid_data, type = "class")

#Evaluation
confusionMatrix(data = predictTREEgini, reference = valid_data_y, positive = "None")


predictTREEinf <- predict(modelTREEinf, valid_data, type = "class")

#Evaluation
confusionMatrix(data = predictTREEinf, reference = valid_data_y, positive = "None")


#WE USE GINI BECAUSE IT WAS THE FIRST ONE COMPUTED - IT IS INDIFERENT SINCE THE RESULTS ARE EQUAL FOR BOTH

summary(modelTREEgini)
rpart.plot(modelTREEgini)

predictTREE <- predict(modelTREEgini, test_data, type = "class")

#Evaluation
confusionMatrix(data = predictTREE, reference = test_data_y, positive = "None")

#####################################################################
#Maybe try pruning (looked at her example from class today)
printcp(modelTREEgini)
bestcp <- modelTREEgini$cptable[which.min(modelTREEgini$cptable[,"xerror"]), "CP"]
plotcp(modelTREEgini)

#Prune the tree - It is not good, but we can show this in the report
modelTREEgini_pruned <- prune(modelTREEgini, cp = bestcp)
predictTREEgini_pruned <- predict(modelTREEgini_pruned, test_data, type = "class")

#Evaluation
confusionMatrix(data = predictTREEgini_pruned, reference = test_data_y, positive = "None")

```
 

## Random Forest

```{r, echo=TRUE}

# train_data_x_rf = train_data_x[,]
# train_data_x_rf$ORIGIN_AIRPORT <- NULL
# train_data_x_rf$DESTINATION_AIRPORT <- NULL
# train_data_x_rf$DAY<- NULL
# 
# test_data_rf = test_data[,]
# test_data_rf$ORIGIN_AIRPORT <- NULL
# test_data_rf$DESTINATION_AIRPORT <- NULL
# test_data_rf$DAY <- NULL
# 
# modelRF = randomForest(train_data_y~., data = train_data_x, mtry = 3, importance = TRUE)
# #Cannot handle variables with more than 53 categories... so we had to try to reduce the data set just to see if this will work
# 
# 
# predictRF <- predict(modelRF, test_data_rf)
# 
# #varImpPlot(modelRF)
# 
# #Confusion matrix
# table(predictRF, test_data_y)
# 
# #Compute the TREE error
# mean(test_data_y != predictRF)
# 
# #Compute TREE accuracy
# mean(test_data_y == predictRF) 

```

## Support Vector Classifier
```{r, echo=TRUE}
# modelSVM = svm(train_data_y~., data = train_data_x, kernel = "linear", cost = 10)
# 
# predictSVM <- predict(modelSVM, test_data)
# 
# #Confusion matrix
# table(predictSVM, test_data_y)
# 
# #Compute the TREE error
# mean(test_data_y != predictSVM)
# 
# #Compute TREE accuracy
# mean(test_data_y == predictSVM) 
#   
# #Find the best airline using the histogram
# tableSVM <- cbind(test_data_x, predictSVM)
# 
# 
# #Barplot of the delays per Airline  
# barplot(table(tableSVM$predictSVM, tableSVM$AIRLINE), main="Categories of delay by Airline",
#   xlab="Airlines", col=c("red", "blue"), args.legend = list(x="topleft"),
#   legend = rownames(table(tableSVM$predictSVM, tableSVM$AIRLINE)), beside=TRUE)
# 
# #Get the percentages of delayed and not delayed for Airline
# countdelaySVM <- as.matrix(table(tableSVM$predictSVM, tableSVM$AIRLINE))
# countdelaySVM <- sweep(countdelaySVM, 2, colSums(countdelaySVM), FUN = '/')
# 
# colnames(countdelaySVM) = c("AA","AS", "B6","DL","EV","F9","HA","MQ","NK","OO","UA","US","VX","WN") 
# rownames(countdelaySVM) = c("Delayed", "None")
# 
# #Get the most and least delayed Airline
# mostdelayedSVM <- colnames(countdelaySVM)[which.is.max(countdelaySVM[1,])] 
# mostdelayedSVM      #Airline which had the higher value of delay
# 
# leastdelayedSVM <- colnames(countdelaySVM)[which.is.max(countdelaySVM[2,])] 
# leastdelayedSVM      #Airline which had the lower value of delay
# 
# #See sorted list of delayed and non delayed
# mostdelayed_sortSVM <- sort(countdelaySVM[1,], decreasing = TRUE) #Higher percentage means high delay
# leastdelayed_sortSVM <- sort(countdelaySVM[2,], decreasing = TRUE) #Higher percentage means low delay
# 
# mostdelayed_sortSVM
# leastdelayed_sortSVM
# ```
# 
# ## Logistic Regression - problems with the presence of categories for all variables in train and test sets
# ```{r, echo=TRUE}
# train_data_logreg = train_data[,]
# train_data_logreg$ORIGIN_AIRPORT <- NULL
# train_data_logreg$DESTINATION_AIRPORT <- NULL
# train_data_logreg$MONTH <- NULL
# train_data_logreg$DAY <- NULL
# train_data_logreg$DAY_OF_WEEK <- NULL
# 
# test_data_logreg = test_data[,]
# test_data_logreg$ORIGIN_AIRPORT <- NULL
# test_data_logreg$DESTINATION_AIRPORT <- NULL
# test_data_logreg$MONTH <- NULL
# test_data_logreg$DAY <- NULL
# test_data_logreg$DAY_OF_WEEK <- NULL
# 
# delay = vector(length = length(train_data_y))
# delay = ifelse(train_data_y == "None",0,1)
# #We use family = binomial since our response is of type factor
# 
# #Fit model
# modelLR <- glm(formula = formula(delay~.), data = train_data, family = "binomial", control = list(maxit = 5000))
# 
# coef(modelLR)
# modelLR
# 
# predictLR <- predict(modelLR, test_data_logreg)
# predictLR
# 
# # n = obs
# #round(n*0.15)
# 
# DELAY <- vector(mode = "character",length = 1500)
# for (i in 1:1500) {
# DELAY[i] = ifelse(predictLR[i] < 30, "Delayed", "None")
# }
# 
# #Confusion matrix
# table(DELAY, test_data_y)
# 
# #Compute the TREE error
# mean(test_data_y != predictLR)
# 
# #Compute TREE accuracy
# mean(test_data_y == predictLR)
```



### Multinomial Logistic regression

   
```{r , echo=TRUE}
#Fit model
# modelMLR <- multinom(formula = formula(train_data_y~.), data= train_data)
# predictMLR <- predict(modelMLR, test_data_x)
# 
# #Confusion matrix
# table(predictMLR, test_data_y)
# 
# #Compute MLR error
# mean(test_data_y != predictMLR)
# 
# #Compute MLR accuracy
# mean(test_data_y == predictMLR)
```

## KNN - not able to consider factor and continuous variables simultaneously

```{r , echo=FALSE}
# co = rainbow_hcl(3)
# 
# #Make model
# ks = 1:10 #Choose K from 1 to 10
# classification_errors = sapply(ks,function(k) {
#   pred.knn <- kknn(formula = formula(train_data_y~.), train = train_data, test = valid_data, k = k, distance = 1)
#   fit <- fitted(pred.knn)
#   mean(valid_data_y != fit)   #I think this is the accuracy, but it should be the classification variables
# })
# 
# #Choose the model that has the lowest classification error, use this K
# k_optimal = which.min(classification_errors)   #The best k is 7
# 
# #Plot errors as a function of k
# plot(ks, classification_errors, type="o", pch = 16, ylim = c(0, 1), col = co[2],
#      xlab = "Number of neighbors", ylab="Misclassification error")
# 
# pred.knn.m1 <- kknn(formula = formula(train_data_y~.), train = train_data, test = test_data, k = k_optimal, distance = 1)
# 
# predictKNN <- predict(pred.knn.m1, test = test_data_x)
# 
# #Classification error
# mean(test_data_y != predictKNN)
# 
# #Accuracy
# mean(test_data_y == predictKNN)
# 
# #Confusion matrix
# table(test_data_y , predictKNN)
```


### Linear Discriminant Analysis
   
```{r , echo=TRUE}
#Fit model
# train_data_lda = train_data[,]
# train_data_lda $ORIGIN_AIRPORT <- NULL
# train_data_lda $DESTINATION_AIRPORT <- NULL
# 
# modelLDA <-lda(formula = formula(train_data_y~.), data = train_data_lda)
# 
# predictLDA <- predict(modelLDA, test_data_x)
# 
# #Confusion matrix
# table(predictLDA, test_data_y)
# 
# #Compute LDA error
# mean(test_data_y != predictLDA)
# 
# #Compute LDA accuracy
# mean(test_data_y == predictLDA)
```
